---
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE,comment = "#>")
```

## Overview

Parametric Face Drawings (PFD; Day and Davidenko, 2019) are schematic, though realistic, parameterized line drawings of faces based on the statistical distribution of human facial feature. Principal components analysis applied to a set of 400 landmarks manually placed on a demographically diverse sample yields a "face space" that can be used for face modeling applications. This R package is a partial port of the Matlab code of Day and Davidenko hosted at the Open Science Framework: https://osf.io/6uds5/

Below we demonstrate some of the basic features of the R package.

### Generate and plot a random face from a demographically diverse face space

```{r}
library(pcfacespace)
face <- pcfacespace::gen_face()
plot(pfd_splines(face))
```

### Morph two random faces

```{r}

normcoef <- function(x, type="F") {
  x/norm(as.matrix(x), "F")
}


distinctiveness <- 10

## generate 166 normally distributed princpal component coefficients for face 1
face1 <- normcoef(rnorm(166))
## generate 166 normally distributed princpal component coefficients for face 2
face2 <- normcoef(rnorm(166))

## create a "spline" representation of face1, face2, and the average of face1 and face2 (the morph).
f1 <- pfd_splines(gen_face(coef=face1*distinctiveness))
f2 <- pfd_splines(gen_face(coef=face2*distinctiveness))
f3 <- pfd_splines(gen_face(coef=normcoef(face1+face2)*distinctiveness))

multiplot(list(f1,f2,f3), nrow=1, ncol=3)

```

### Push two faces away from each other in PC-space using a "face difference"

```{r}


## generate 166 normally distributed princpal component coefficients for face 1
face1 <- normcoef(rnorm(166)) 
## generate 166 normally distributed princpal component coefficients for face 2
face2 <- normcoef(rnorm(166))

## create a "face difference" vector
delta <- face1 - face2

## push the first face away from the second face by 1 difference units.
delta_face1 = pcfacespace::gen_face(1.2*(distinctiveness) * normcoef(delta+face1))

## push the second face away from the first face by 1 difference units (notice the sign reversal).
delta_face2 = pcfacespace::gen_face(1.2*(distinctiveness) * normcoef(-delta+face2))

multiplot(list(pfd_splines(gen_face(coef=face1*distinctiveness)),
               pfd_splines(gen_face(coef=face2*distinctiveness)),
               pfd_splines(delta_face1),pfd_splines(delta_face2)), nrow=2, ncol=2)




```

```{r}

## push the first face away from the second face by .5 difference units.
delta_face1 = pcfacespace::gen_face(.5*delta + face1)

## push the second face away from the first face by .5 difference units (notice the sign reversal).
delta_face2 = pcfacespace::gen_face(-.5*delta + face2)

multiplot(list(pfd_splines(gen_face(coef=face1)),pfd_splines(gen_face(coef=face2)),
               pfd_splines(delta_face1),pfd_splines(delta_face2)), nrow=2, ncol=2)

```

```{r}

## push the first face away from the second face by .3 difference units.
delta_face1 = pcfacespace::gen_face(.3*delta + face1)

## push the second face away from the first face by .3 difference units (notice the sign reversal).
delta_face2 = pcfacespace::gen_face(-.3*delta + face2)

multiplot(list(pfd_splines(gen_face(coef=face1)),pfd_splines(gen_face(coef=face2)),
               pfd_splines(delta_face1),pfd_splines(delta_face2)), nrow=2, ncol=2)

```


## References

Day, J., & Davidenko, N. (2019). Parametric face drawings: A demographically diverse and customizable face space model. Journal of vision, 19(11), 7-7.


